\huge The human visual system preserves the hierarchy of 2-dimensional pattern regularity

### 2021-02-20 Writeup/kohler_clarke_2021.tex
Fixed citations and increased linespacing

@@

Symmetries are abundant in natural and man-made environments, due to a complex interplay of physical forces that govern pattern formation in nature. Symmetrical patterns have been created and appreciated by human cultures throughout history and since the gestalt movement of the early 20th century, symmetry has been recognized as important for visual perception. Symmetry contributes to the perception of shapes \cite{RN1311,RN1682}, scenes \cite{RN1824} and surface properties \cite{RN1166}, as well as the social process of mate selection \cite{RN1337}. Most of this work has focused on mirror symmetry or \textit{reflection}, with much less attention being paid to the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection}. In the two spatial dimensions relevant for images, these four symmetries can be combined in 17 distinct ways, \textit{the wallpaper groups} \cite{RN1562,RN1563,RN1425}. Previous work on a subset of four of the wallpaper groups used functional MRI to demonstrate that rotation symmetries in wallpapers elicit parametric responses in several areas in occipital cortex, beginning with visual area V3 \cite{RN1725}. This effect was also robust with electroencephalography (EEG), whether measured using Steady-State Visual Evoked Potentials (SSVEPs)\cite{RN1725} or event-related paradigms \cite{RN1959}. Here we extend this work by collecting SSVEPs and psychophysical data from human participants viewing the full set of wallpaper groups. We measure responses in visual cortex to 16 out of the 17 wallpaper groups, with the 17th serving as a control stimulus. Our goal is to provide a more complete picture of how wallpaper groups are represented in the human visual system.

A wallpaper group is a topologically discrete group of isometries of the Euclidean plane, i.e. transformations that preserve distance \cite{RN1425}. Wallpaper groups differ in the number and kind of these transformations. In mathematical group theory, when the elements of one group is completely contained in another, the inner group is called a subgroup of the outer group \cite{RN1425}. The full list of subgroup relationships is listed in Section 1.4.2 of the Supplementary Material. Subgroup relationships between wallpaper groups can be distinguished by their indices. The index of a subgroup relationship is the number of cosets, i.e. the number of times the subgroup is found in the supergroup \cite{RN1425}. As an example, let us consider groups \textit{P2} and \textit{P6}. If we ignore the translations in two directions that both groups share, group \textit{P6} consists of the set of rotations \{$0^{\circ}$, $60^{\circ}$, $120^{\circ}$, $180^{\circ}$, $240^{\circ}$, $300^{\circ}$\}, in which \textit{P2} \{$0^{\circ}$, $180^{\circ}$\} is contained. \textit{P2} is thus a subgroup of \textit{P6}, and \textit{P6} can be generated by combining \textit{P2} with rotations \{$0^{\circ}$, $120^{\circ}$, $240^{\circ}$\}. Because \textit{P2} is repeated three times in \textit{P6}, \textit{P2} is a subgroup of \textit{P6} with index 3 \cite{RN1425}. The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \cite{RN1711}.

@@

The two datasets we present here make it possible to assess the extent to which both behavior and brain responses follow the hierachy of complexity expressed by the subgroup relationships. Based on previous brain imaging work showing that patterns with more axes of symmetry produce greater activity in visual cortex \cite{RN1170, RN1959, RN1725, keefe_2018}, we hypothesized that more complex groups would produce larger SSVEPs. For the psychophysical data, we hypothesized that more complex groups would lead to shorter symmetry detection thresholds, based on previous data showing that under a fixed presentation time, discriminability increases with the number of symmetry axes in the pattern \cite{wagemans_1991}. Our results confirm both hypotheses, and show that activity in human visual cortex is remarkably consistent with the hierarchical relationships between the wallpaper groups, with SSVEP amplitudes and psychophysical thresholds following these relationships at a level that is far beyond chance. The human visual system thus appears to encode all of the fundamental symmetries using a representational structure that closely approximates the subgroup relationships from group theory.

@@

The stimuli used in our two experiments were generated from random-noise textures, which made it possible to generate multiple exemplars from each of the wallpaper groups, as described in detail elsewhere \cite{RN1725}. We generated control stimuli matched to each exemplar in the main stimulus set, by scrambling the phase but maintaining the power spectrum. All wallpaper groups are inherently periodic because of their repeating lattice structure. Phase scrambling maintains this periodicity so the phase-scrambled control images all belong to group \textit{P1} regardless of group membership of the original exemplar. \textit{P1} contains no symmetries other than translation, while all other groups contain translation in combination with one or more of the other three fundamental symmetries (reflection, rotation, glide reflection) \cite{RN1425}. In our SSVEP experiment, this stimulus set allowed us to isolate brain activity specific to the symmetry structure in the exemplar images from activity associated with modulation of low-level features, by alternating exemplar images and control exemplars. In this design, responses to structural features beyond the shared power spectrum, including any symmetries other than translation, are isolated in the odd harmonics of the image update frequency \cite{RN1725,RN1540, RN1954}. Thus, the combined magnitude of the odd harmonic response components can be used as a measure of the overall strength of the visual cortex response.

@@

SSVEP data from four of the wallpaper groups (\textit{P2}, \textit{P3}, \textit{P4} and \textit{P6}) was previously published as part of our earlier demonstration of parametric responses to rotation symmetry in wallpaper groups\cite{RN1725}. We replicate that result using our Bayesian approach, and find an analogous parametric effect in the psychophysical data (see Supplementary Figure 4.1). We also conducted an analysis testing for an effect of index in our two datasets, and found that subgroup relationships with higher indices tended to produce greater pairwise differences between the subgroup and supergroup, for both SSVEP RMS and symmetry detection thresholds (see Supplementary Figure 4.2). The effect of index is relatively weak, but the fact that there is a measurable index effect can nonetheless be taken as preliminary evidence that representations of symmetries in wallpaper groups may be compositional.

@@

Here we show that beyond merely responding to the elementary symmetry operations of reflection \cite{RN1170} and rotation \cite{RN1725}, the visual system represents the hierarchical structure of the 17 wallpaper groups, and thus every composition of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures. Both SSVEP amplitudes and symmetry detection thresholds preserve the hierarchy of complexity among the wallpaper groups that is captured by the subgroup relationships \cite{RN1711}. For the SSVEP, this remarkable consistency was specific to the odd harmonics of the stimulus frequency that are known to capture the symmetry-specific response \cite{RN1725} and to electrodes in a region-of-interest (ROI) over occipital cortex. When the same analysis was done using the odd harmonics from electrodes over parietal cortex (Supplementary Figure 3.3) or even harmonics from electrodes over occipital cortex (Supplementary Figure 3.4), the data was substantially less consistent with the subgroup relationships (yellow and green lines, Figure 4).

The current data provide a description of the visual system’s response to the complete set of symmetries in the two-dimensional plane. Our design precludes us from independently measure the response to \textit{P1}, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 1), we can conclude that the relationships between P1 and all other groups, where \textit{P1} is the subgroup, are also preserved by the visual system. The subgroup relationships are in many cases not obvious perceptually, and most participants had no knowledge of group theory. Thus, the visual system’s ability to preserve the subgroup hierarchy does not depend on explicit knowledge of the relationships. Previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \cite{RN1253}, they tend to sort them into fewer groups than there actually are (4-12 groups) and it is common for exemplars from different wallpaper groups to be sorted in the same group \cite{RN172}. The more controlled two-interval forced choice approach used in the current behavioral experiment allows us to show that more granular representations of wallpaper groups are measurable in behavior.

@@

The correspondence between responses in the visual system and group theory that we demonstrate here, may reflect a form of implicit learning that depends on the structure of the natural world. The environment is itself constrained by physical forces underlying pattern formation and these forces are subject to multiple symmetry constraints \cite{RN1634}. The ordered structure of responses to wallpaper groups could be driven by a central tenet of neural coding, that of efficiency. If coding is to be efficient, neural resources should be distributed to capture the structure of the environment with minimum redundancy considering the visual geometric optics, the capabilities of the subsequent neural coding stages and the behavioral goals of the organism \cite{RN1758, RN1760, RN1757, RN1756}. Early work within the efficient coding framework suggested that natural images had a $1/f$ spectrum and that the corresponding redundancy between pixels in natural images could be coded efficiently with a sparse set of oriented filter responses, such as those present in the early visual pathway \cite{RN1740, RN1446}. Our results suggest that the principle of efficient coding extends to a much higher level of structural redundancy – that of symmetries in visual images.

The 17 wallpaper groups are completely regular, and relatively rare in the visual environment, especially when considering distortions due to perspective and occlusion. Near-regular textures, however, abound in the visual world, and can be approximated as deformed versions of the wallpaper groups \cite{RN1519}. The correspondence between visual cortex responses and group theory demonstrated here may indicate that the visual system represents visual textures using a similar scheme, with the wallpaper groups serving as anchor points in representational space. This framework resembles norm-based encoding strategies that have been proposed for other stimulus classes, most notably faces \cite{RN435}, and leads to the prediction that adaptation to wallpaper patterns should distort perception of near-regular textures, similar to the aftereffects found for faces \cite{RN1768}. Field biologists have demonstrated that animals respond more strongly to exaggerated versions of a learned stimulus, referred to as “supernormal” stimuli \cite{RN1775}. In the norm-based encoding framework, wallpaper groups can be considered \textit{supertextures}, exaggerated examples of the near-regular textures that surround us. Artists may consciously or unconsciously create supernormal stimuli, to capture the essence of the subject and evoke strong responses in the audience \cite{RN1764}. Wallpaper groups are visually compelling and symmetries have been widely used in human artistic expression going back to the Neolithic age \cite{RN1949}. If wallpapers are in fact supertextures, this prevalence may be a direct result of the strategy the human visual system has adopted for texture encoding.

@@

\subsection*{Stimulus Generation}
Exemplars from the different wallpaper groups were generated using a modified version of the methodology developed by Clarke and colleagues\cite{RN172} that we have described in detail elsewhere\cite{RN1725}. Briefly, exemplar patterns for each group were generated from random-noise textures, which were then repeated and transformed to cover the plane, according to the symmetry axes and geometric lattice specific to each group. The use of noise textures as the starting point for stimulus generation allowed the creation of an almost infinite number of distinct exemplars of each wallpaper group. To make individual exemplars as similar as possible we replaced the power spectrum of each exemplar with the median across exemplars within a group. We then generated control exemplars that had the same power spectrum as the exemplar images by randomizing the phase of each exemplar image. The phase scrambling eliminates rotation, reflection and glide-reflection symmetries within each exemplar, but the phase-scrambled images inherent the spectral periodicity arising from the periodic tiling. This means that all control exemplars, regardless of which wallpaper group they are derived from, are transformed into another symmetry group, namely \textit{P1}. \textit{P1} is the simplest of the wallpaper groups, and contains only translations of a region whose shape derives from the lattice. Because the different wallpaper groups have different lattices, \textit{P1} controls matched to different groups have different power spectra. Our experimental design takes these differences into account by comparing the neural responses evoked by each wallpaper group to responses evoked by the matched control exemplars.

\subsection*{Stimulus Presentation}
Stimulus Presentation. For the EEG experiment, the stimuli were shown on a 24.5" Sony Trimaster EL PVM-2541 organic light emitting diode (OLED) display at a screen resolution of $1920\times1080$ pixels, 8-bit color depth and a refresh rate of 60 Hz, viewed at a distance of 70 cm. The mean luminance was 69.93 cd/m2 and contrast was 95\%. The diameter of the circular aperture in which the wallpaper pattern appeared was $13.8^\circ$ of visual angle presented against a mean luminance gray background. Stimulus presentation was controlled using in-house software. For the psychophysics experiment, the stimuli were shown on a $48 \times 27$cm VIEWPixx/3D LCD Display monitor, model VPX-VPX-2005C, resolution $1920 \times 1080$ pixels, with a viewing distance of approximately 40cm and linear gamma. Stimulus presentation was controlled using MatLab and Psychtoolbox-3 \cite{kleiner2007,brainard1997spatial}. The diameter of the circular aperture for the stimuli was $21.5^\circ$.

@@

\subsection*{Psychophysics Procedure}
The experiment consisted of 16 blocks, one for each of the wallpaper groups (excluding \textit{P1}). We used a two-interval forced choice approach. In each trial, participants were presented with two stimuli (one of which was the wallpaper group for the current block of trials, the other being \textit{P1}), one after the other (inter-stimulus interval of 700ms). After each stimuli had been presented, it was masked with white noise for 300ms. After both stimuli had been presented, participants made a response on the keyboard to indicate whether they thought the first or second image contained more symmetry. Each block started with 10 practice trials, (stimulus display duration of 500ms) to allow participants to familiarise themselves with the current block's wallpaper pattern. If they achieved an accuracy of 9/10 in these trials they progressed to the rest of the block, otherwise they carried out another set of 10 practise trials. This process was repeated until the required accuracy of 9/10 was obtained. The rest of the block consisted of four interleaved staircases (using the QUEST algorithm \cite{watson1983quest}, full details given in the SI) of 30 trials each. On average, a block of trials took around 10 minutes to complete.

@@

\subsection*{SSVEP Analysis}
Response waveforms were generated for each group by selective filtering in the frequency domain. For each participant, the average Fourier coefficients from the two sessions were averaged over trials and sessions. The SSVEP paradigm we used allowed us to separate symmetry-related responses from non-specific contrast transient responses. Previous work has demonstrated that symmetry-related responses are predominantly found in the odd harmonics of the stimulus frequency, whereas the even harmonics consist mainly of responses unrelated to symmetry, that arise from the contrast change associated with the appearance of the second image \cite{RN1954, RN1725}. This functional distinction of the harmonics allowed us to generate a single-cycle waveform containing the response specific to symmetry, by filtering out the even harmonics in the spectral domain, and then back-transforming the remaining signal, consisting only of odd harmonics, into the time-domain. For our main analysis, we averaged the odd harmonic single-cycle waveforms within a six-electrode region of interest (ROI) over occipital cortex (electrodes 70, 74, 75, 81, 82, 83). These waveforms, averaged over participants, are shown in Figure 1. The same analysis was done for the even harmonics and for the odd harmonics within a six electrode ROI over parietal cortex (electrodes 53, 54, 61, 78, 79, 86; see Supplementary Figure 1.1). The root-mean square values of these waveforms, for each individual participant, were used to determine whether each of the wallpaper subgroup relationships were preserved in the brain data.

\subsection*{Defining the list of subgroup relationships}
In order to get the complete list of subgroup relationships, we digitized Table 4 from Coxeter \cite{RN1711} (shown in Supplementary Table 1.2). After removing identity relationships (i.e. each group is a subgroup of itself) and the three pairs of wallpapers groups that are subgroups of each other (e.g. \textit{PM} is a subgroup of \textit{CM}, and \textit{CM} is a subgroup of \textit{PM}) we were left with a total of 63 unambiguous subgroups that were included in our analysis.

\subsection*{Bayesian Analysis of SSVEP and Psychophysical data}
Bayesian analysis was carried out using R (v3.6.1) \cite{R} with the \texttt{brms} package (v2.9.0) \cite{burkner2017} and rStan (v2.19.2 \cite{rStan}). The data from each experiment were modelled using a Bayesian generalised mixed effect model with wallpaper group being treated as a 16 level factor, and random effects for participant. The SSVEP data and symmetry detection threshold durations were modelled using log-normal distributions with weakly informative, $ \mathcal{N}(0, 2)$, priors. After fitting the model to the data, samples were drawn from the posterior distribution of the two datasets, for each wallpaper group. These samples were then recombined to calculate the distribution of differences for each of the 63 pairs of subgroup and supergroup. These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | data)$. For further technical details, please see the Supplementary Materials at \url{https://osf.io/f3ex8/} where the full R code, model specification, prior and posterior predicitive checks, and model diagnoastics, can be found.

>>>>

\renewcommand{\baselinestretch}{1.3}

@@

Symmetries are abundant in natural and man-made environments, due to a complex interplay of physical forces that govern pattern formation in nature. Symmetrical patterns have been created and appreciated by human cultures throughout history and since the gestalt movement of the early 20th century, symmetry has been recognized as important for visual perception. Symmetry contributes to the perception of shapes \citep{RN1311,RN1682}, scenes \citep{RN1824} and surface properties \citep{RN1166}, as well as the social process of mate selection \citep{RN1337}. Most of this work has focused on mirror symmetry or \textit{reflection}, with much less attention being paid to the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection}. In the two spatial dimensions relevant for images, these four symmetries can be combined in 17 distinct ways, \textit{the wallpaper groups} \citep{RN1562,RN1563,RN1425}. Previous work on a subset of four of the wallpaper groups used functional MRI to demonstrate that rotation symmetries in wallpapers elicit parametric responses in several areas in occipital cortex, beginning with visual area V3 \citep{RN1725}. This effect was also robust with electroencephalography (EEG), whether measured using Steady-State Visual Evoked Potentials (SSVEPs)\citep{RN1725} or event-related paradigms \citep{RN1959}. Here we extend this work by collecting SSVEPs and psychophysical data from human participants viewing the full set of wallpaper groups. We measure responses in visual cortex to 16 out of the 17 wallpaper groups, with the 17th serving as a control stimulus. Our goal is to provide a more complete picture of how wallpaper groups are represented in the human visual system.

A wallpaper group is a topologically discrete group of isometries of the Euclidean plane, i.e. transformations that preserve distance \citep{RN1425}. Wallpaper groups differ in the number and kind of these transformations. In mathematical group theory, when the elements of one group is completely contained in another, the inner group is called a subgroup of the outer group \citep{RN1425}. The full list of subgroup relationships is listed in Section 1.4.2 of the Supplementary Material. Subgroup relationships between wallpaper groups can be distinguished by their indices. The index of a subgroup relationship is the number of cosets, i.e. the number of times the subgroup is found in the supergroup \citep{RN1425}. As an example, let us consider groups \textit{P2} and \textit{P6}. If we ignore the translations in two directions that both groups share, group \textit{P6} consists of the set of rotations \{$0^{\circ}$, $60^{\circ}$, $120^{\circ}$, $180^{\circ}$, $240^{\circ}$, $300^{\circ}$\}, in which \textit{P2} \{$0^{\circ}$, $180^{\circ}$\} is contained. \textit{P2} is thus a subgroup of \textit{P6}, and \textit{P6} can be generated by combining \textit{P2} with rotations \{$0^{\circ}$, $120^{\circ}$, $240^{\circ}$\}. Because \textit{P2} is repeated three times in \textit{P6}, \textit{P2} is a subgroup of \textit{P6} with index 3 \citep{RN1425}. The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \citep{RN1711}.

@@

The two datasets we present here make it possible to assess the extent to which both behavior and brain responses follow the hierachy of complexity expressed by the subgroup relationships. Based on previous brain imaging work showing that patterns with more axes of symmetry produce greater activity in visual cortex \citep{RN1170, RN1959, RN1725, keefe_2018}, we hypothesized that more complex groups would produce larger SSVEPs. For the psychophysical data, we hypothesized that more complex groups would lead to shorter symmetry detection thresholds, based on previous data showing that under a fixed presentation time, discriminability increases with the number of symmetry axes in the pattern \citep{wagemans_1991}. Our results confirm both hypotheses, and show that activity in human visual cortex is remarkably consistent with the hierarchical relationships between the wallpaper groups, with SSVEP amplitudes and psychophysical thresholds following these relationships at a level that is far beyond chance. The human visual system thus appears to encode all of the fundamental symmetries using a representational structure that closely approximates the subgroup relationships from group theory.

@@

The stimuli used in our two experiments were generated from random-noise textures, which made it possible to generate multiple exemplars from each of the wallpaper groups, as described in detail elsewhere \citep{RN1725}. We generated control stimuli matched to each exemplar in the main stimulus set, by scrambling the phase but maintaining the power spectrum. All wallpaper groups are inherently periodic because of their repeating lattice structure. Phase scrambling maintains this periodicity so the phase-scrambled control images all belong to group \textit{P1} regardless of group membership of the original exemplar. \textit{P1} contains no symmetries other than translation, while all other groups contain translation in combination with one or more of the other three fundamental symmetries (reflection, rotation, glide reflection) \citep{RN1425}. In our SSVEP experiment, this stimulus set allowed us to isolate brain activity specific to the symmetry structure in the exemplar images from activity associated with modulation of low-level features, by alternating exemplar images and control exemplars. In this design, responses to structural features beyond the shared power spectrum, including any symmetries other than translation, are isolated in the odd harmonics of the image update frequency \citep{RN1725,RN1540, RN1954}. Thus, the combined magnitude of the odd harmonic response components can be used as a measure of the overall strength of the visual cortex response.

@@

SSVEP data from four of the wallpaper groups (\textit{P2}, \textit{P3}, \textit{P4} and \textit{P6}) was previously published as part of our earlier demonstration of parametric responses to rotation symmetry in wallpaper groups\citep{RN1725}. We replicate that result using our Bayesian approach, and find an analogous parametric effect in the psychophysical data (see Supplementary Figure 4.1). We also conducted an analysis testing for an effect of index in our two datasets, and found that subgroup relationships with higher indices tended to produce greater pairwise differences between the subgroup and supergroup, for both SSVEP RMS and symmetry detection thresholds (see Supplementary Figure 4.2). The effect of index is relatively weak, but the fact that there is a measurable index effect can nonetheless be taken as preliminary evidence that representations of symmetries in wallpaper groups may be compositional.

@@

Here we show that beyond merely responding to the elementary symmetry operations of reflection \citep{RN1170} and rotation \citep{RN1725}, the visual system represents the hierarchical structure of the 17 wallpaper groups, and thus every composition of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures. Both SSVEP amplitudes and symmetry detection thresholds preserve the hierarchy of complexity among the wallpaper groups that is captured by the subgroup relationships \citep{RN1711}. For the SSVEP, this remarkable consistency was specific to the odd harmonics of the stimulus frequency that are known to capture the symmetry-specific response \citep{RN1725} and to electrodes in a region-of-interest (ROI) over occipital cortex. When the same analysis was done using the odd harmonics from electrodes over parietal cortex (Supplementary Figure 3.3) or even harmonics from electrodes over occipital cortex (Supplementary Figure 3.4), the data was substantially less consistent with the subgroup relationships (yellow and green lines, Figure 4).

The current data provide a description of the visual system’s response to the complete set of symmetries in the two-dimensional plane. Our design precludes us from independently measure the response to \textit{P1}, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 1), we can conclude that the relationships between P1 and all other groups, where \textit{P1} is the subgroup, are also preserved by the visual system. The subgroup relationships are in many cases not obvious perceptually, and most participants had no knowledge of group theory. Thus, the visual system’s ability to preserve the subgroup hierarchy does not depend on explicit knowledge of the relationships. Previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \citep{RN1253}, they tend to sort them into fewer groups than there actually are (4-12 groups) and it is common for exemplars from different wallpaper groups to be sorted in the same group \citep{RN172}. The more controlled two-interval forced choice approach used in the current behavioral experiment allows us to show that more granular representations of wallpaper groups are measurable in behavior.

@@

The correspondence between responses in the visual system and group theory that we demonstrate here, may reflect a form of implicit learning that depends on the structure of the natural world. The environment is itself constrained by physical forces underlying pattern formation and these forces are subject to multiple symmetry constraints \citep{RN1634}. The ordered structure of responses to wallpaper groups could be driven by a central tenet of neural coding, that of efficiency. If coding is to be efficient, neural resources should be distributed to capture the structure of the environment with minimum redundancy considering the visual geometric optics, the capabilities of the subsequent neural coding stages and the behavioral goals of the organism \citep{RN1758, RN1760, RN1757, RN1756}. Early work within the efficient coding framework suggested that natural images had a $1/f$ spectrum and that the corresponding redundancy between pixels in natural images could be coded efficiently with a sparse set of oriented filter responses, such as those present in the early visual pathway \citep{RN1740, RN1446}. Our results suggest that the principle of efficient coding extends to a much higher level of structural redundancy – that of symmetries in visual images.

The 17 wallpaper groups are completely regular, and relatively rare in the visual environment, especially when considering distortions due to perspective and occlusion. Near-regular textures, however, abound in the visual world, and can be approximated as deformed versions of the wallpaper groups \citep{RN1519}. The correspondence between visual cortex responses and group theory demonstrated here may indicate that the visual system represents visual textures using a similar scheme, with the wallpaper groups serving as anchor points in representational space. This framework resembles norm-based encoding strategies that have been proposed for other stimulus classes, most notably faces \citep{RN435}, and leads to the prediction that adaptation to wallpaper patterns should distort perception of near-regular textures, similar to the aftereffects found for faces \citep{RN1768}. Field biologists have demonstrated that animals respond more strongly to exaggerated versions of a learned stimulus, referred to as “supernormal” stimuli \citep{RN1775}. In the norm-based encoding framework, wallpaper groups can be considered \textit{supertextures}, exaggerated examples of the near-regular textures that surround us. Artists may consciously or unconsciously create supernormal stimuli, to capture the essence of the subject and evoke strong responses in the audience \citep{RN1764}. Wallpaper groups are visually compelling and symmetries have been widely used in human artistic expression going back to the Neolithic age \citep{RN1949}. If wallpapers are in fact supertextures, this prevalence may be a direct result of the strategy the human visual system has adopted for texture encoding.

@@

\subsection*{Stimulus Generation}
Exemplars from the different wallpaper groups were generated using a modified version of the methodology developed by Clarke and colleagues\citep{RN172} that we have described in detail elsewhere\citep{RN1725}. Briefly, exemplar patterns for each group were generated from random-noise textures, which were then repeated and transformed to cover the plane, according to the symmetry axes and geometric lattice specific to each group. The use of noise textures as the starting point for stimulus generation allowed the creation of an almost infinite number of distinct exemplars of each wallpaper group. To make individual exemplars as similar as possible we replaced the power spectrum of each exemplar with the median across exemplars within a group. We then generated control exemplars that had the same power spectrum as the exemplar images by randomizing the phase of each exemplar image. The phase scrambling eliminates rotation, reflection and glide-reflection symmetries within each exemplar, but the phase-scrambled images inherent the spectral periodicity arising from the periodic tiling. This means that all control exemplars, regardless of which wallpaper group they are derived from, are transformed into another symmetry group, namely \textit{P1}. \textit{P1} is the simplest of the wallpaper groups, and contains only translations of a region whose shape derives from the lattice. Because the different wallpaper groups have different lattices, \textit{P1} controls matched to different groups have different power spectra. Our experimental design takes these differences into account by comparing the neural responses evoked by each wallpaper group to responses evoked by the matched control exemplars.

\subsection*{Stimulus Presentation}
Stimulus Presentation. For the EEG experiment, the stimuli were shown on a 24.5" Sony Trimaster EL PVM-2541 organic light emitting diode (OLED) display at a screen resolution of $1920\times1080$ pixels, 8-bit color depth and a refresh rate of 60 Hz, viewed at a distance of 70 cm. The mean luminance was 69.93 cd/m2 and contrast was 95\%. The diameter of the circular aperture in which the wallpaper pattern appeared was $13.8^\circ$ of visual angle presented against a mean luminance gray background. Stimulus presentation was controlled using in-house software. For the psychophysics experiment, the stimuli were shown on a $48 \times 27$cm VIEWPixx/3D LCD Display monitor, model VPX-VPX-2005C, resolution $1920 \times 1080$ pixels, with a viewing distance of approximately 40cm and linear gamma. Stimulus presentation was controlled using MatLab and Psychtoolbox-3 \citep{kleiner2007,brainard1997spatial}. The diameter of the circular aperture for the stimuli was $21.5^\circ$.

@@

\subsection*{Psychophysics Procedure}
The experiment consisted of 16 blocks, one for each of the wallpaper groups (excluding \textit{P1}). We used a two-interval forced choice approach. In each trial, participants were presented with two stimuli (one of which was the wallpaper group for the current block of trials, the other being \textit{P1}), one after the other (inter-stimulus interval of 700ms). After each stimuli had been presented, it was masked with white noise for 300ms. After both stimuli had been presented, participants made a response on the keyboard to indicate whether they thought the first or second image contained more symmetry. Each block started with 10 practice trials, (stimulus display duration of 500ms) to allow participants to familiarise themselves with the current block's wallpaper pattern. If they achieved an accuracy of 9/10 in these trials they progressed to the rest of the block, otherwise they carried out another set of 10 practise trials. This process was repeated until the required accuracy of 9/10 was obtained. The rest of the block consisted of four interleaved staircases (using the QUEST algorithm \citep{watson1983quest}, full details given in the SI) of 30 trials each. On average, a block of trials took around 10 minutes to complete.

@@

\subsection*{SSVEP Analysis}
Response waveforms were generated for each group by selective filtering in the frequency domain. For each participant, the average Fourier coefficients from the two sessions were averaged over trials and sessions. The SSVEP paradigm we used allowed us to separate symmetry-related responses from non-specific contrast transient responses. Previous work has demonstrated that symmetry-related responses are predominantly found in the odd harmonics of the stimulus frequency, whereas the even harmonics consist mainly of responses unrelated to symmetry, that arise from the contrast change associated with the appearance of the second image \citep{RN1954, RN1725}. This functional distinction of the harmonics allowed us to generate a single-cycle waveform containing the response specific to symmetry, by filtering out the even harmonics in the spectral domain, and then back-transforming the remaining signal, consisting only of odd harmonics, into the time-domain. For our main analysis, we averaged the odd harmonic single-cycle waveforms within a six-electrode region of interest (ROI) over occipital cortex (electrodes 70, 74, 75, 81, 82, 83). These waveforms, averaged over participants, are shown in Figure 1. The same analysis was done for the even harmonics and for the odd harmonics within a six electrode ROI over parietal cortex (electrodes 53, 54, 61, 78, 79, 86; see Supplementary Figure 1.1). The root-mean square values of these waveforms, for each individual participant, were used to determine whether each of the wallpaper subgroup relationships were preserved in the brain data.

\subsection*{Defining the list of subgroup relationships}
In order to get the complete list of subgroup relationships, we digitized Table 4 from Coxeter \citep{RN1711} (shown in Supplementary Table 1.2). After removing identity relationships (i.e. each group is a subgroup of itself) and the three pairs of wallpapers groups that are subgroups of each other (e.g. \textit{PM} is a subgroup of \textit{CM}, and \textit{CM} is a subgroup of \textit{PM}) we were left with a total of 63 unambiguous subgroups that were included in our analysis.

\subsection*{Bayesian Analysis of SSVEP and Psychophysical data}
Bayesian analysis was carried out using R (v3.6.1) \citep{R} with the \texttt{brms} package (v2.9.0) \citep{burkner2017} and rStan (v2.19.2 \citep{rStan}). The data from each experiment were modelled using a Bayesian generalised mixed effect model with wallpaper group being treated as a 16 level factor, and random effects for participant. The SSVEP data and symmetry detection threshold durations were modelled using log-normal distributions with weakly informative, $ \mathcal{N}(0, 2)$, priors. After fitting the model to the data, samples were drawn from the posterior distribution of the two datasets, for each wallpaper group. These samples were then recombined to calculate the distribution of differences for each of the 63 pairs of subgroup and supergroup. These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | data)$. For further technical details, please see the Supplementary Materials at \url{https://osf.io/f3ex8/} where the full R code, model specification, prior and posterior predicitive checks, and model diagnoastics, can be found.
